\usefont{OT1}{bch}{b}{n} \normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt] \huge Deep learning transcriptomic model for prediction of pan-cancer chemotherapeutic sensitivity \\

### 2020-06-01 manuscript/manuscript.tex
edited abstract

% Abstract
\begin{abstract}
	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes genes expression profiles to stratify tumours into similar pan-drug sensitivity profiles. This development would allow for identification of candidate drugs for treatment that maximize response and minimize cytotoxic burden. As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning architecture. Pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups. Due to the high dimensional nature of the microarray data, biological agnostic feature selection was conducted to highlight genes with the highest predictive value. We developed and tuned a deep learning neural net architecture to predict response groups. Cell line clusters showed a significant difference in the therapeutic response, generated from IC50 values of all available chemotherapeutics. The feature space was reduced to 300 genes, which functional profiling indicated enriched primarily the focal adhesion, ECM-receptor and proteoglycan interaction pathways. Hyperparameter tuning of the deep learning model dictated a 4 hidden layer, diamond shaped architecture with a predictive accuracy of 91.7\%. \vspace{1em}

>>>>

% Abstract
\begin{abstract}
	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles. This development would allow for identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden. As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model. A pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups. Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity. Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value. The feature space was reduced to 300 genes, which functional profiling indicated enriched primarily the focal adhesion, ECM-receptor and proteoglycan interaction pathways. Using these selected genes, a deep learning neural network architecture was developed to predict response groups. Hyperparameter tuning of the deep learning model dictated a 4 hidden layer architecture with a predictive accuracy of 91.7\%. \vspace{1em}
